Aligning marine species range data to better serve science and conservation

### 2016-08-28 ms_2_methods.Rmd
some updates to methods and res/disc re: corals and aquamaps clipping

To examine errors of commission related to depth, we selected corals as a case study due to their importance in supporting biodiversity as well as their dependence on photosynthesis.  We created a 200 m bathymetry raster from Natural Earth's 200 m bathymetry polygon [@natural_earth_XXX] ___(Fig. SXXX)___ and masked our IUCN coral rasters to identify mapped coral presence below 200 m.  The resulting maps were again compared to the AquaMaps rasters to examine distribution alignment and area ratio.

>>>>

To examine errors of commission related to depth, we selected corals as a case study due to their importance in supporting biodiversity as well as their dependence on photosynthesis.  Extracting data from the IUCN API [@REF?] we identified the depth limitations of each of the coral species mapped in the IUCN dataset.  We created a 200 m bathymetry raster from Natural Earth's 200 m bathymetry polygon [@natural_earth_XXX] ___(Fig. SXXX)___ and masked our IUCN coral rasters to identify mapped coral presence below 200 m.  The resulting maps were again compared to the AquaMaps rasters to examine distribution alignment and area ratio.

AquaMaps uses Food and Agriculture Organization of the United Nations (FAO) Major Fishing Area [@REF] boundaries as a rough method of constraining species ranges to appropriate georegions.  To explore the impact of these artificial constraints on species range predictions, we chose to focus on boundaries defined by longitude, as they seem less likely to relate to ecological conditions than boundaries defined by latitude.  As an example, we identified all AquaMaps species whose eastern or western range limit between between 5° S and 25° N latitudes coincided with the large vertical boundary between FAO regions 71 and 77 at 175° W longitude.

### 2016-08-28 ms_3_res_disc_concl.Rmd
some updates to methods and res/disc re: corals and aquamaps clipping

[results] Fig. 3A shows the 463 coral species mapped in the IUCN dataset, with their ranges broken into proportional area deeper and shallower than 200 m.  ___show a map of all coral species aggregated, with deeper than 200 m cells shown___  None of these species is indicated to occur deeper than 200 m, and 94% are confined to waters shallower than 50 m; seven of the mapped species had no reported depth information.  Clipping coral ranges to shallower than 200 m eliminated an average of 47.6% of the total predicted area while still allowing for a generous estimate of suitable habitat.

@@

[results] In constraining coral ranges to shallow waters, we see a strong increase in the apparent alignment of species maps between IUCN and AquaMaps.  Using the original quadrant definitions from Fig. 2B, we see in Fig. 3B and 3C a massive shift in 354 paired map species due to constraining coral depths to 200 m.  Membership in the "well-aligned" quadrant jumped from 22.4% to 76.2%, with a corresponding decrease in all other quadrants.

[discussion] We cannot know for certain the true distribution of each of these corals; much of the apparent over-prediction of coral ranges is likely due to experts taking a conservative and precautionary approach.  But a sensible shift in method drastically decreases the likelihood of introducing commission errors, with little chance of introducing omission errors, greatly improving our confidence in the remaining reported ranges.  Note that this change applies just as well to the IUCN coral maps that are not included in the paired map analysis, and likely to other reef-associated flora and fauna.  While species depth preferences are an easy and consistent means of constraining range predictions, other conditions such as salinity and temperature could be cautiously used to refine the results of expert opinions.

***What is still missing here is something to show that AquaMaps ranges, clipped to FAO Major Fishing Areas, are probably introducing errors of commission as well*** - _the fish don't see that FAO boundary and suddenly decide not to cross; most obvious in the boundaries defined by longitudes - sudden vertical edges in range.  Can we show this with a couple of example maps?  That wasn't received well for the corals, but I can't think of a clear and unambiguous way to analyze this problem._

>>>>

[results] Fig. 3A shows the 463 coral species mapped in the IUCN dataset, with their ranges broken into proportional area deeper and shallower than 200 m.  According to IUCN descriptions, none of these species is indicated to occur deeper than 200 m, and 94% are confined to waters shallower than 50 m; seven of the mapped species had no reported depth information.  Clipping coral ranges to shallower than 200 m eliminated an average of 47.6% of the total predicted area while still allowing for a generous estimate of suitable habitat.

@@

[results] In constraining coral ranges to shallow waters, we see a strong increase in the apparent alignment of species maps between IUCN and AquaMaps.  Using the original quadrant definitions from Fig. 2B, we see in Fig. 3B a massive shift in 354 paired-map coral species due to constraining coral depths to 200 m.  Membership in the "well-aligned" quadrant jumped from 22.4% to 76.2%, with a corresponding decrease in all other quadrants.  See SXXX to examine the shifts of individual species among the quadrants.

[discussion] We cannot know for certain the true distribution of each of these corals; much of the apparent over-expansion of coral ranges is likely due to experts taking a conservative and precautionary approach.  But a sensible shift in method drastically decreases the likelihood of introducing commission errors, with little chance of introducing omission errors, greatly improving our confidence in the remaining reported ranges.  Note that this change applies just as well to the IUCN coral maps that are not included in the paired map analysis, and likely to other reef-associated flora and fauna.  While species depth preferences are an easy and consistent means of constraining range predictions, other conditions such as salinity and temperature could be cautiously used to refine the results of expert opinions.

## Georegion constraint exploration

- AM clipped map

[results]
